String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-05-08 2023-1-19/sections/introduction.tex
related work

\paragraph*{Related work} \zhilin{to be cleaned} To handle the problems in software verification reasoning about strings and integers, string theories with regular membership predicate and linear length constraints are raised. String solvers such as Z3Str3\cite{z3str3}, Z3Seq\cite{z3seq}, BD+23\cite{BD+23}, CVC4\cite{cvc4}, CVC5\cite{cvc5}, (Z3-)Trau\cite{trau}\cite{z3trau}, Ostrich(+)\cite{ostrich}\cite{atva2020}, Sloth\cite{sloth}, S3\cite{s3}, and ABC\cite{abc} are developed on these theories. Although string theory is easily
undecidable\cite{undecidable_1}\cite{undecidable_2}, the satisfiability problems for string constraints of regular expressions, linear integer arithmetic, and the string length is decidable\cite{theory_BD+23}. There are mainly two strategies for solving string constraints: (1) DPLL(T)-based\cite{dpll_t} string solvers\cite{z3str3}\cite{cvc5} apply heuristic derivation rules to unwind the regular expression gradually and identify several classes of simplification techniques for efficiency. (2) Automata-based string solvers\cite{trau}\cite{z3trau}\cite{ostrich}\cite{atva2020}\cite{BD+23} construct a finite automaton to model the regular expression and use the emptiness problem of the automaton (under linear integer arithmetic) to decide the satisfiability of string constraints.

>>>>

\paragraph*{Related work} 
We discuss more related work about regular expressions with counting and real-world regular expressions in programming languages here.  
Determinism of regular expressions with counting was investigated in~\cite{GGM12,CL15}. Counting-set automata was proposed in~\cite{redos_lenka,HS+23} to enable fast matching of a subclass of regular expressions with counting.  Moreover, a variant of nondeterministic counter automata, called bit vector automata, was proposed recently in \cite{GKM23} to enable fast matching of a more expressive class of regular expressions with counting.   Nevertheless, nonemptiness problem of these automata, which is the key for automata-based string constraint solving, was not considered in ~\cite{redos_lenka,HS+23,GKM23}.
Real-world regular expressions in programming languages include some features beyond classical regular expressions, e.g. the greedy/lazy Kleene star, capturing groups, and back references. Real-world regular expressions haven been addressed in symbolic execution of Javascript programs~\cite{LMK19} and string constraint solving \cite{CF+22}.

%To handle the problems in software verification reasoning about strings and integers, string theories with regular membership predicate and linear length constraints are raised. String solvers such as Z3Str3\cite{z3str3}, Z3Seq\cite{z3seq}, BD+23\cite{BD+23}, CVC4\cite{cvc4}, CVC5\cite{cvc5}, (Z3-)Trau\cite{trau}\cite{z3trau}, Ostrich(+)\cite{ostrich}\cite{atva2020}, Sloth\cite{sloth}, S3\cite{s3}, and ABC\cite{abc} are developed on these theories. Although string theory is easily
%undecidable\cite{undecidable_1}\cite{undecidable_2}, the satisfiability problems for string constraints of regular expressions, linear integer arithmetic, and the string length is decidable\cite{theory_BD+23}. There are mainly two strategies for solving string constraints: (1) DPLL(T)-based\cite{dpll_t} string solvers\cite{z3str3}\cite{cvc5} apply heuristic derivation rules to unwind the regular expression gradually and identify several classes of simplification techniques for efficiency. (2) Automata-based string solvers\cite{trau}\cite{z3trau}\cite{ostrich}\cite{atva2020}\cite{BD+23} construct a finite automaton to model the regular expression and use the emptiness problem of the automaton (under linear integer arithmetic) to decide the satisfiability of string constraints.
